An Ethnography of Online Performance Artists

### 2017-03-17 proposals/Brown_Magic_Grant_2017/main.tex
proposal coming close to 'complete' status

@@

\lhead{} % controls the left corner of the header
% \chead{} % controls the center of the header
\rhead{Ali Alkhatib} % controls the right corner of the header
\lfoot{last updated} % controls the left corner of the footer
\cfoot{} % controls the center of the footer
\rfoot{Page~\thepage} % controls the right corner of the footer
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}

\thispagestyle{empty}

@@

% \newlist{inlinelist}{enumerate*}{1}
% \setlist*[inlinelist,1]{
%   label=\arabic*),
% }
\makeatletter

\def\maketitle{%
\par{\centering {\scshape \huge \textbf{\@title}}%
\par{\@author}
\par{DRAFT \DTMcurrenttime~\today}%
% \par{DRAFT \DTMcurrenttime~on~\today}%
\par}}
% \smallskip \hrule \smallskip }

\definecolor{PineGreen}{HTML}{008800}
\definecolor{Red}{HTML}{FF0000}
\definecolor{Blue}{HTML}{0000FF}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}

\title{Ethnographic Study of Online Entertainers}
\author{Ali Alkhatib}
\date{\today}

\definecolor{Blue}{HTML}{0000FF}
\newcommand{\topic}[1]{{\color{Blue}#1}}
% \renewcommand{\topic}[1]{{#1}}
\definecolor{Gray}{HTML}{A0A0A0}
\newcommand{\ugh}[1]{{\color{Gray}#1}}
\newenvironment{blah}{\par\color{Gray}\itshape}{\par}

@@

\topic{One issue frustrating this potential research area is that
the platforms have changed so dramatically in as few as 5 to 10 years that
it can seem as though researchers are trying to wade through shifting sands.}
Scholarship on content creators loses bearing rapidly when economic factors motivate aspiring professional entertainers to join the community~\cite{Hamilton:2014:STF:2611105.2557048};
the tension between these entertainers experimenting with content remixing and
the holders of the copyright of the source material
(for example, see~\cite{Hilderbrand48})
shifts dramatically when
the platform's copyright enforcement mechanisms advance
not just incrementally but substantially~\cite{kim2012institutionalization}.

% \section*{Insight}
\topic{We recognize that \textit{historical analysis}
% Studying contemporary phenomena through an established framing
can be a helpful way of
making sense of contemporary phenomena.}
As we found in a recent paper framing crowd work and gig work as a modern instantiation of \textit{piecework},
scholarship from a parallel or similar domain can
provide us some grounding to make sense of what we've seen, as well as inform predictions~\cite{pieceworkCrowdworkGigwork}.
% This positioning of an existing subject in historical context not only
% provided us with guidance toward some of the major threads of research
% in crowdsourcing, but
% also informed some of the potential futures of crowdsourcing and
% the ever growing gig work labor markets.

@@

\topic{This connection has purchase, but it has flaws, mostly owing to the unique nature of ``networked publics''~\cite{boyd2007youth}.}
For one thing, most video sharing websites keep videos
% --- the modern performances ---
in perpetuity
--- that is, unless a complaint over copyright infringement is filed, in which case the lifespan of that ``performance'' can be cut short.
The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
YouTube comments linger for years after they're made.
And these say nothing of the different nature of algorithmic enforcement of rules and policies, and
the near--perfect ability of these platforms to redirect revenue and even people's focus instantaneously.

@@

\section*{Research Plan}
\ali{Basic plan:
\begin{itemize}
  \item contact YouTube channels that have more than 1M subscribers
        (and have mentioned YouTube policies at some point
          (suggesting maybe passing interest in discussing meta topics)?)
  \item Reach out by any contact methods they provide (if none, then skip)
  \item Semi--structured interview guiding towards:
  \begin{itemize}
    \item interaction with platform \& its designers
    \item engagement with audience
    \item coordination with other YouTubers (if any)
  \end{itemize}
  \item tentative questions include:
  \begin{itemize}
    \item platform questions
    \begin{enumerate}
      \item when did you switch to (streaming/YouTubing) as primary career?
      \item what's YouTube's relationship with you?
      \item if you were telling a new YouTuber how to get started,
            what would you say they should know about YouTube
            (the system, the company, etc\dots)
            that's not immediately obvious?
    \end{enumerate}
    \item audience questions
    \begin{enumerate}
      \item how do you interact with audiences?
      \item what sorts of comments do you get?
      \item do you reply?
      \item do you change your future content based on feedback in the comments?
    \end{enumerate}
    \item peer questions
    \begin{enumerate}
      \item what's your sense of the community as a whole?
      \item how many other YouTubers do you know
            (let's say you've had a chat with them in the last month and
             will probably do so again in the next month)?
      \item are you all cooperative? competitive? what's the culture like?
    \end{enumerate}
  \end{itemize}
\end{itemize}
}

@@

\pagebreak
\printbibliography{}
\end{document}

>>>>

\usepackage{caption}
\captionsetup[table]{labelformat=empty}

@@

\lhead{} % controls the left corner of the header
% \chead{} % controls the center of the header
\rhead{Alkhatib \& Nagy} % controls the right corner of the header
\lfoot{last updated} % controls the left corner of the footer
\cfoot{} % controls the center of the footer
\rfoot{Page~\thepage} % controls the right corner of the footer
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}

\AtBeginBibliography{\small}
\thispagestyle{empty}

@@

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}
\makeatletter

\def\maketitle{%
\par{\centering {\scshape \LARGE \textbf{\@title}}%
\par{\@author}
\par{DRAFT \DTMcurrenttime~\today}%
% \par{DRAFT \DTMcurrenttime~on~\today}%
\par}}
% \smallskip \hrule \smallskip }

\definecolor{PineGreen}{HTML}{008800}
\definecolor{Red}{HTML}{FF0000}
\definecolor{Blue}{HTML}{0000FF}
\definecolor{Black}{HTML}{000000}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}

\title{An Ethnography of Online Performance Artists}
\author{Ali Alkhatib \& Jeffrey Nagy}
\date{\today}
\urlstyle{same}
\hypersetup{
  colorlinks,
  citecolor=Black,
  urlcolor=Blue}

\definecolor{Blue}{HTML}{0000FF}
\newcommand{\topic}[1]{{\color{Blue}#1}}
\renewcommand{\topic}[1]{{#1}}
\definecolor{Gray}{HTML}{A0A0A0}
\newcommand{\ugh}[1]{{\color{Gray}#1}}
\newenvironment{blah}{\par\color{Gray}\itshape}{\par}

@@

\topic{One issue frustrating this potential research area is that
the platforms have changed so dramatically in as few as 5 to 10 years that
it can seem as though researchers are trying to wade through shifting sands.}
Scholarship on content creators loses bearing rapidly when
economic factors motivate aspiring professional entertainers to join the community~\cite{Hamilton:2014:STF:2611105.2557048};
the tension between these entertainers experimenting with content remixing and
the holders of the copyright of the source material
(for example, see~\cite{Hilderbrand48})
shifts dramatically when
the platform's copyright enforcement mechanisms advance
not just incrementally but substantially~\cite{kim2012institutionalization}.
And the popularity of candidate field sites can vary chaotically
(see, for example, Vine reaching as many as 200 million active users and
abrupt shutdown less than a year later~\cite{vineDecline}).

% \section*{Insight}
\topic{We recognize that \textit{historical analysis}
% Studying contemporary phenomena through an established framing
can be a helpful way of
making sense of contemporary phenomena.}
As we found in a recent paper framing crowd work and gig work as a modern instantiation of \textit{piecework},
scholarship from a parallel or similar domain can
provide us some grounding to make sense of what we've seen, as well as inform predictions~\cite{pieceworkCrowdworkGigwork}.

@@

\topic{This connection has purchase, but it has gaps, mostly owing to the unique nature of the internet~\cite{miller2011understanding}.}
For one thing, most video sharing websites keep videos
in perpetuity
--- that is, unless a complaint over copyright infringement is filed, in which case the lifespan of that ``performance'' can be cut short.
The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
YouTube comments linger for years after they're made, forming in some sense ``networked publics''~\cite{boyd2007youth}.
And these say nothing of the different nature of algorithmic enforcement of rules and policies, and
the near--perfect ability of these platforms to redirect revenue and even people's focus instantaneously.

\topic{We propose to fill in these gaps in our framing of online performance as a modern instantiation of street performance.}
How does the \textit{permanence} of digitally archived videos affect people's willingness to experiment with new formats and styles,
and \textit{how can we enable and encourage riskier, potentially more rewarding experimentation}?
How do networked publics influence whether and how performers engage with their audiences,
and \textit{how can we design publics that engage the audience with each other and with the performer}?
How do laws, and the algorithmically instantiated policies, on video sharing and streaming platforms spur or stifle creative remixing?

@@

\section*{Research Plan}
\topic{We propose to carry out qualitative research with major online content producers
on streaming platforms such as Twitch and YouTube, and video sharing sites like YouTube.}
Much of this work will consist of ethnographic fieldwork, semi--structured interviews, and
open--ended exploration of the field site as we find parallels and divergences in the domain
of online performance art from street and other emergent formats of performance and entertainment.
% With that being said, we can generally frame our agenda in five components
% ---
% \begin{inlinelist}
%   \item Surveying the landscape of online performance platforms,
%         identifying performers who may be willing to serve as ``informants'', and
%         recruiting interviewing participants;
%   \item Investigating the relationships and tensions performers have with \textit{audiences};
%   \item Exploring the relationships performers have with \textit{other performers};
%   \item Studying the relationships performers have with \textit{the platforms themselves}; and finally
%   \item Designing interventions to facilitate creative experimentation in this nascent format,
%         informed largely by each of the earlier facets of online performance.
% \end{inlinelist}
A rough timeline of these phases can be described as such:

\begin{tabular}{p{2cm}|p{13cm}}
  \textbf{Term} & \textbf{Plan} \\
\hline
Q1  & Identify potential participants and begin recruiting participants;
      begin preliminary interviews and fieldwork.
      Research specifically focusing on performers' \textit{relationships with audiences}
      and noting areas of frustration for performers and intervention opportunities \\
\hline
Q2  & Qualitative fieldwork toward understanding performers' \textit{relationships with other performers},
      with some attention toward the potential for performer affiliations \\
\hline
Q3  & Fieldwork looking at performers' \textit{relationships with their respective platforms}
      (for example: YouTube, Twitch) \\
\hline
Q4  & Exploring and iterating on \textit{design interventions} with YouTubers and streamers
      to advance values discussed earlier
      (audience engagement, online performer affiliation, etc\dots)  \\
\end{tabular}

\topic{Our interview process will }
We will use semi--structured interviews
to remain open to participants guiding our research agenda toward
whatever they consider to be compelling topics.
With that being said

% \ali{Basic plan:
% \begin{itemize}
%   \item contact YouTube channels that have more than 1M subscribers
%         (and have mentioned YouTube policies at some point
%           (suggesting maybe passing interest in discussing meta topics)?)
%   \item Reach out by any contact methods they provide (if none, then skip)
%   \item Semi--structured interview guiding towards:
%   \begin{itemize}
%     \item interaction with platform \& its designers
%     \item engagement with audience
%     \item coordination with other YouTubers (if any)
%   \end{itemize}
%   \item tentative questions include:
%   \begin{itemize}
%     \item platform questions
%     \begin{enumerate}
%       \item when did you switch to (streaming/YouTubing) as primary career?
%       \item what's YouTube's relationship with you?
%       \item if you were telling a new YouTuber how to get started,
%             what would you say they should know about YouTube
%             (the system, the company, etc\dots)
%             that's not immediately obvious?
%     \end{enumerate}
%     \item audience questions
%     \begin{enumerate}
%       \item how do you interact with audiences?
%       \item what sorts of comments do you get?
%       \item do you reply?
%       \item do you change your future content based on feedback in the comments?
%     \end{enumerate}
%     \item peer questions
%     \begin{enumerate}
%       \item what's your sense of the community as a whole?
%       \item how many other YouTubers do you know
%             (let's say you've had a chat with them in the last month and
%              will probably do so again in the next month)?
%       \item are you all cooperative? competitive? what's the culture like?
%     \end{enumerate}
%   \end{itemize}
% \end{itemize}
% }

@@

% \pagebreak
% \renewcommand\bibname{{\small References}}
\setlength\bibitemsep{0pt}

\renewcommand*{\bibfont}{\tiny}

\printbibliography{}
\end{document}
